\section*{Task 3: Authentication}
The third and last task of the lab was about authentication, and the exercises were of a theoretical rather than practical nature.
\subsection*{Exercises:}
\begin{enumerate}
\item % (a)
Hashing a list of passwords with a one-way hash function, say $f$, rather than having the passwords stored in cleartext is a method for reducing the impact of a database leak. Authentication given a password $p$ in such a system is done by computing the hash $f(p)$ of the password and comparing it with the hash for the specific user stored in the database.

In the event that the password hashes are exposed to an attacker, there is no trivial way for him or her to retrieve the passwords corresponding to the compromised password hashes, by the definition of a one-way hash function. Thus, there is no trivial way of authenticating despite having access to both the hash function and the hashes. If an attacker finds the passwords stored in cleartext, on the other hand, he or she can authenticate as any of the affected users on the system. Furthermore, because passwords are frequently reused in more than one system, the attacker might be able to authenticate falsely on other systems as well. In other words, the impact of a cleartext password leak could encompass more than just the attacked system.

While there is no \emph{trivial} way of finding a password corresponding to a specific hash value, there is a cumbersome but in practice feasible way of doing so: password guessing. The idea is to repeatedly come up with a guess for the password, hash it, and compare the result with the hash value in question, continuing until a password guess with the specific hash is found. The hash values for guesses for common passwords can be precomputed into a so-called rainbow table that allow for these passwords to be looked up given their hash value.

Salting is a method for making password guessing more difficult. In addition to storing a hash value for each user, a random string $s$---the salt---is stored for each user as well. Authentication given a password $p$ is done by computing the hash of a combination of the password and the salt, such as $s + p$ where ``$+$'' denotes a concatenation. The resulting hash, $f(s + p)$, is then compared with the hash value in the database.

If an attacker gets hold of the database, which contains a salt and hash for each user, and aims for finding the password of \emph{any} of the users, the addition of the salt will make password guessing more difficult. Let $n$ be the number of users. Under the assumption that each user has a unique salt, the attacker will have to compute $n$ hashes for each password guess, compared to the single computation that is required with hashing alone. Salting offers increased security also against attacks with precomputed tables, seeing as one table would be needed for each possible salt in order to cover the same passwords as without salting.
\item % (b)
Assuming that the salt and salted hash are known to an attacker, salting may make the attack on a specific account more difficult, depending on whether precomputed tables are used.

If precomputed tables are used and the attacker did not choose the specific user because the attacker had a precomputed table for that specific salt, salting will make the attack more difficult in that more precomputed tables will be required.

Without precomputation, however, each password guess will require one computation of a hash value, assuming that the hash is computed like mentioned before: $f(s + g)$ for hash function $f$, salt $s$, password guess $g$, and concatenation operator $+$. The only difference in the case that salting is not used is that no salt has to be concatenated; the hash function still has to be evaluated once for each guess. Under the assumption that concatenating the string takes negligible time compared to evaluating the hash function, salting does not make the attack in question harder.
\item \highergradesonly
  \begin{itemize}
    \item
    \item
  \end{itemize}
\item
  \begin{itemize}
    \item 
    \item
  \end{itemize}
\end{enumerate}
